---
title: "Catalase_analysis.Rmd"
author: "Sam Gurr"
date: "2024-10-31"
output: html_document
---

### SET UP

* note: add your own directory below to navigate to RAnalysis folder containing the data, output, and script subfolders 
that reflect the google drive. alterntively clone the repository EAD-ASEB-Cvirginica_Thermal_Performance from github
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# SET WORKING DIRECTORY 
# knitr::opts_knit$set(root.dir = "C:/Users/katherine.mcfarland/Documents/GitHub/EAD-ASEB-Cvirginica_Thermal_Performance/larvae") # Katie's
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/Cgigas_SMIF_2024/RAnalysis") # Sam's
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/EAD-ASEB-Cvirginica_Thermal_Performance/RAnalysis") # Sam's work

```


## Load packages 

* use install.packges('pacakge name') if you do not have these downloaded
```{r setup, include=TRUE}

library(ggpubr)
library(ggplot2)
library(dplyr)
library(lmtest) # to receive p value from betareg model
# library(FSA) # for the Dun test post hoc for SRH non-parametric 2 way anova]
library(emmeans)
library(tidyverse)
library(car)
library(lmerTest)
library(tidyr)
library(reshape2)
library(ggpubr)
library(nlme)
library(rcompanion) # to run the Schrier -Ray-Hare non parametric 2 way 
library(ggpmisc) # stat_poly for inserting equation and R2 for ggplot line 
```




```{r load experiment metadata and edit}

# objective to relate unique smaple IDs to treatment(s)
# the tric being to change the tube id in this raw format to match that in the plate metadata to merge later.. dont worry its easy!
exp.metadata <- read.csv('Data/experiment_metadata/SMIF_Sample_Info_SG_ParallelAnalyses.csv', sep = ",", header=T) %>% 
                      select(c('Set', 'Tube.ID', 'Assay..SOD.CAT.')) %>%  # call target columns 
                      filter(Assay..SOD.CAT. %in% 'SOD') %>%  # redundant sample ids for each assay, arbitrarily choose SOD to get unique IDs
                      mutate(Time = gsub("_.*", "", Tube.ID), # make column Time by calling everying before the first _
                             ID   = gsub("_", "", gsub("_S_B$", "", Tube.ID)),
                             Treatment = case_when(
                                                  grepl("Control", Set) ~ "control",
                                                  grepl("OA", Set) ~ "OA",
                                                  grepl("HW", Set) ~ "hypoxia",
                                                  TRUE ~ "baseline"
                             )) %>% 
                      select(ID, Time, Treatment)


```
## Function to load spec files

-   At your service, from 'spec to table' ...I'm so funny

```{r Spec_to_table}

Spec_to_table <- function(datapath, platenumber, daterecorded) {
  
# import the data in raw text form exported from the spectrophotometer
        rawtextfile <- read.delim(datapath,
                              header = F,
                              skip=3,
                              fileEncoding="latin1")
  

        rawmatrix   <- as.matrix(rawtextfile[c(1:8),c(3:14)])
        colnames(rawmatrix) = c("1","2","3","4","5","6","7","8","9","10","11","12")
        rownames(rawmatrix) = c("A","B","C","D","E","F","G","H")
        
        raw_table <- as.data.frame.table(rawmatrix, responseName = "value") %>% 
                                    dplyr::rename(well_row=Var1, well_column=Var2, absorbance_nm=value) %>% 
                                    dplyr::mutate(well=paste0(well_row,well_column),
                                                  Run_date =daterecorded,
                                                  Plate=platenumber) %>% 
                                    dplyr::select(-c(well_row,well_column))
        # reorder the columns because I prefer it this way..
        
        raw_table_ordered <- raw_table[,c(3,4,2,1)]
}


```

-   Use the spec_to_table to import raw spec txt files and compile to a table, yes that is what it does

```{r load Catalase raw data and merge}

# plate 1 
catalase_plate1 <- Spec_to_table('Data/raw_data/catalase/20250219_catalase_plate1.txt', 1, 20250219) %>%  
                                  dplyr::rename(Cat_Absorbance_560nm=absorbance_nm) 
# plate 2 
catalase_plate2 <- Spec_to_table('Data/raw_data/catalase/20250219_catalase_plate2.txt', 2, 20250219) %>%  
                                  dplyr::rename(Cat_Absorbance_560nm=absorbance_nm) 

# master 
catalase_raw <- rbind(catalase_plate1, catalase_plate2)

```

```{r load Total protein correction data for catalase}

# plate 1 
totalprotein_plate1 <- Spec_to_table('Data/raw_data/catalase/20250221_total_protein_catalase_plate1.txt', 1, 20250221) %>%  
                                  dplyr::rename(TP_Absorbance_562nm=absorbance_nm) 

# plate 2
totalprotein_plate2 <- Spec_to_table('Data/raw_data/catalase/20250221_total_protein_catalase_plate2.txt', 2, 20250221) %>%  
                                  dplyr::rename(TP_Absorbance_562nm=absorbance_nm) 

# master 
totalprotein_raw <- rbind(totalprotein_plate1, totalprotein_plate2)

```


```{r load reference metadata and call target datasets}

metadata <- read.csv('Data/raw_data/plate_metadata.csv', sep = ",", header=T) 

# rename a few columns so they can merge with the data (in next chunk)

# Plate 1 
meta.catalase_plate1 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Catalase" & Plate.Number == 1) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)
# Plate 2
meta.catalase_plate2 <- metadata %>% 
                        dplyr::filter(Assay.Type %in% "Catalase" & Plate.Number == 2) %>% 
                        dplyr::rename(Plate = Plate.Number,
                                      well  = Well.ID)



# Plate 1 
meta.totalprotein_plate1 <- metadata %>% 
                                  dplyr::filter(Assay.Type %in% "Total protein - Catalase" & Plate.Number == 1) %>% 
                                  dplyr::rename(Plate = Plate.Number,
                                                well  = Well.ID)
# Plate 2
meta.totalprotein_plate2 <- metadata %>% 
                                  dplyr::filter(Assay.Type %in% "Total protein - Catalase" & Plate.Number == 2) %>% 
                                  dplyr::rename(Plate = Plate.Number,
                                                well  = Well.ID)
```


* merge metadata with data 
```{r merge metadata with data}

# Catalase
# Plate 1 
catalase_plate1_master <- merge(catalase_plate1, meta.catalase_plate1, 
                                by = c('Plate', 'well')) 

nrow(meta.catalase_plate1) == nrow(catalase_plate1_master) # must be TRUE
catalase_plate1_master <- subset(catalase_plate1_master, !is.na(Sample.Type)) # now 96  total rows with NAs omitted (not sample, blank nor stnadard)


# Plate 2
catalase_plate2_master <- merge(catalase_plate2, meta.catalase_plate2, 
                                by = c('Plate', 'well')) 

nrow(meta.catalase_plate2) == nrow(catalase_plate2_master) # must be TRUE
catalase_plate2_master <- subset(catalase_plate2_master, !is.na(Sample.Type)) # now 52  total rows with NAs omitted (not sample, blank nor stnadard)






# Total protein
# Plate 1 
totalprotein_plate1_master <- merge(totalprotein_plate1, meta.totalprotein_plate1, 
                                by = c('Plate', 'well')) 

nrow(meta.totalprotein_plate1) == nrow(totalprotein_plate1_master) # must be TRUE
totalprotein_plate1_master <- subset(totalprotein_plate1_master, !is.na(Sample.Type)) # now 96  total rows with NAs omitted (not sample, blank nor stnadard)

# Plate 2 
totalprotein_plate2_master <- merge(totalprotein_plate2, meta.totalprotein_plate2, 
                                by = c('Plate', 'well')) 

nrow(meta.totalprotein_plate2) == nrow(totalprotein_plate2_master) # must be TRUE
totalprotein_plate2_master <- subset(totalprotein_plate2_master, !is.na(Sample.Type)) # now 58  total rows with NAs omitted (not sample, blank nor stnadard)

```


* subset standards, blanks, and samples
```{r subset data types}

# Catalase 
# Plate 1 
catalase_plate1.samples   <- catalase_plate1_master %>% dplyr::filter(Sample.Type %in% 'sample') 
catalase_plate1.standards <- catalase_plate1_master %>% dplyr::filter(Sample.Type %in% 'standard') %>% 
                                dplyr::mutate(catalase_UmL = case_when(
                                                                        Standards %in% 'S1' ~ 5,
                                                                        Standards %in% 'S2' ~ 2.5,
                                                                        Standards %in% 'S3' ~ 1.25,
                                                                        Standards %in% 'S4' ~ 0.625,
                                                                        Standards %in% 'S5' ~ 0.313,
                                                                        Standards %in% 'S6' ~ 0.156,
                                                                        Standards %in% 'S7' ~ 0
                                ))

(nrow(catalase_plate1.samples) + 
      nrow(catalase_plate1.standards)) == nrow(catalase_plate1_master) # sanity check, must be true


# Plate 2
catalase_plate2.samples   <- catalase_plate2_master %>% dplyr::filter(Sample.Type %in% 'sample') 
catalase_plate2.standards <- catalase_plate2_master %>% dplyr::filter(Sample.Type %in% 'standard') %>% 
                                dplyr::mutate(catalase_UmL = case_when(
                                                                        Standards %in% 'S1' ~ 5,
                                                                        Standards %in% 'S2' ~ 2.5,
                                                                        Standards %in% 'S3' ~ 1.25,
                                                                        Standards %in% 'S4' ~ 0.625,
                                                                        Standards %in% 'S5' ~ 0.313,
                                                                        Standards %in% 'S6' ~ 0.156,
                                                                        Standards %in% 'S7' ~ 0
                                ))

(nrow(catalase_plate2.samples) + 
      nrow(catalase_plate2.standards)) == nrow(catalase_plate2_master) # sanity check, must be true





# Total protein
# Plate 1 
totalprotein_plate1.samples    <- totalprotein_plate1_master %>% dplyr::filter(Sample.Type %in% 'sample')
totalprotein_plate1.blanks     <- totalprotein_plate1_master %>% dplyr::filter(Sample.Type %in% 'blank')
totalprotein_plate1.standards  <- totalprotein_plate1_master %>% dplyr::filter(Sample.Type %in% c('standard', 'blank'))

(nrow(totalprotein_plate1.samples) + 
      nrow(totalprotein_plate1.standards)) == nrow(totalprotein_plate1_master) # sanity check, must be true


# Plate 2
totalprotein_plate2.samples    <- totalprotein_plate2_master %>% dplyr::filter(Sample.Type %in% 'sample')
totalprotein_plate2.blanks     <- totalprotein_plate2_master %>% dplyr::filter(Sample.Type %in% 'blank')
totalprotein_plate2.standards  <- totalprotein_plate2_master %>% dplyr::filter(Sample.Type %in% c('standard', 'blank'))

(nrow(totalprotein_plate2.samples) + 
      nrow(totalprotein_plate2.standards)) == nrow(totalprotein_plate2_master) # sanity check, must be true

```

## Total protein calculations 

```{r total protein standard curve}

# assign known BCA values to standard IDs A-I
BCA_standards <- rbind(totalprotein_plate1.standards, totalprotein_plate2.standards) %>%  
  dplyr::mutate(BCA_ug_mL = case_when(Standards %in% 'A' ~ 2000,
                                      Standards %in% 'B' ~ 1500,
                                      Standards %in% 'C' ~ 1000,
                                      Standards %in% 'D' ~ 750,
                                      Standards %in% 'E' ~ 500,
                                      Standards %in% 'F' ~ 250,
                                      Standards %in% 'G' ~ 125,
                                      Standards %in% 'H' ~ 25,
                                      Standards %in% 'I' ~ 0)) %>% 
  dplyr::select(Plate, Standards, BCA_ug_mL, TP_Absorbance_562nm)

# Run standard curve, calculate totalprotein 
BCA_background_zero <- BCA_standards %>% 
                        dplyr::filter(Standards %in% 'I') %>% # the zero standard
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# Plate 1, blank to correct by is 0.1289
# Plate 2, blank to correct by is 0.1336



# Absorbance corrected - take the mean of any duplicates
BCA_standards_means <- BCA_standards %>% 
                        dplyr::filter(!BCA_ug_mL %in% 0) %>% 
                        dplyr::mutate(Abs_562nm_cor = 
                                      case_when(Plate == 1 ~ (TP_Absorbance_562nm-0.1289),
                                                Plate == 2 ~ (TP_Absorbance_562nm-0.1336) ) ) %>% 
                        dplyr::select(-TP_Absorbance_562nm) %>% 
                        dplyr::group_by(Plate, Standards, BCA_ug_mL) %>% # group by to get the means
                        dplyr::summarise_each(funs(mean,sd,se=sd(.)/sqrt(n()))) # get all the stats 


# plot it insert the quadratic formaula using ggpmisc
BCA_stand_plots_quadratic <- BCA_standards_means %>% 
                    # QUADRATIC SMOOTH LINE WORKS BEST HERE (MANUFACTURERS INSTRUCTIONS)
                     # dplyr::filter(!(Plate %in% 2 & Standards %in% 'D')) %>% # hash me out to test
                     ggplot(aes(y=mean, x=BCA_ug_mL)) + 
                        geom_point() +
                        theme_bw() +
                        labs(y= "Net Abs 562nm", x = "Protein Concentration in ug/mL") +
                        #geom_line() +
                        #stat_poly_line(color='red') +
                        #geom_smooth() +
                        stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1) +
                        stat_poly_eq(parse=T, aes(label = ..eq.label..), formula=y ~ x + I(x^2)) +
                        ggtitle('Total protein: Quadratic curve') +
                        #stat_poly_eq(use_label(c("eq", "R2"))) +
                        facet_wrap(~Plate) 

# output 
pdf(paste("Output/catalase/plots/Total_protein_standard_curves.pdf", sep =''), 
    width=10, 
    height=7)
print(ggarrange(BCA_stand_plots_quadratic))
dev.off()

```


* Use BCA standard curves to calculate TP 

```{r calc TP}
library(grDevices)
# Standard curve, Plate 1 equation y = -0.0189 + 0.00101x - 7.71x10^-8x^2 - need to solve for x!
# Standard curve, Plate 2 equation y = --0.0189 + 0.00196x - 4.42x10^-7x^2 - need to solve for x!

# Standard curve, Plate 1
a1 <- 4.35*10^-8
b1 <- 0.00171
c1 <- 0.0478
# EQ: (-(b1) + sqrt( (b1^2) - (4*(((a1)-Abs_562nm_cor))*(c1)) ))/(2*a1)

# Standard curve, Plate 2
a2 <- 1.96*10^-8
b2 <- 0.00184
c2 <- 0.0391
# EQ: (-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_cor)) ) ) / (2*a2)


# linear equation plate 1 == (Abs_562nm_cor - 0.192)/0.000993
# linear equation plate 2 == (Abs_562nm_cor - 0.224)/0.000911


# IMPORTANT! we used 25 ul of the standards and 25 ul of the unknowns (samples) 
# therefore we can interpret the unknown direct to the the standard curve without having 
# to account for addition factors, fot example, if we used 5 ul unknown (sample) we would have to adjust 
# by multiplying by 5 to reach the standard curve 

V = 0.025 # 25 ul or 0.025 mL

# Again, remember Plate 1, blank to correct by is 0.1289
# Again, remember Plate 2, blank to correct by is 0.1336

# Sanity check Lets look at the absorbance vs. totla protein concentration data 

TotalProtein_final <- rbind(totalprotein_plate1.samples, totalprotein_plate2.samples) %>% 
                      dplyr::select(Plate,
                                    Sample_ID,
                                    TP_Absorbance_562nm) %>% 
                      # dplyr::filter(!Abs_562nm > 3.99) %>% # data as 4.00 is above the detection limit, omit
                      # dplyr::mutate(Unique_ID = 
                      #                 paste0('Plate:',Plate,' ',
                      #                        'ID:',Sample_ID)) %>% # unique ID t0 group by
                      dplyr::mutate(Abs_562nm_corrected = # correct the raw abs, subtract background
                                   case_when(Plate == 1 ~ (TP_Absorbance_562nm-0.1289), # for plate 1
                                             Plate == 2 ~ (TP_Absorbance_562nm-0.1336) ) ) %>% # for plate 2 
                      dplyr::mutate(TotalProtein_ug_mL = 
                                    case_when(
                                      # linear fr neg discrim. - luckily only two values from plate 2
                                      # Scallop_ID %in% c(33, 51) ~ 
                                      #   ((Abs_562nm_cor - 0.224)/0.000911),
                                      # quadratic for Plate 1
                                      Plate == 1 ~ 
                                        ((-(b1) + sqrt( (b1^2) - (4*a1*(c1-Abs_562nm_corrected)) ) ) / (2*a1)), 
                                      # quadratic for plate 2
                                      Plate == 2 ~ 
                                        ((-(b2) + sqrt( (b2^2) - (4*a2*(c2-Abs_562nm_corrected)) ) ) / (2*a2)) ),
                                    # ug per mL concentration to ug in 25 ul sample 
                                    TotalProtein_ug = TotalProtein_ug_mL*V) %>% 
                      dplyr::mutate(Dilution_factor = 0, 
                                    Volume_ul = 25) %>% 
                      dplyr::mutate(Abs_562nm_raw = TP_Absorbance_562nm) %>% 
                      dplyr::select(Plate, Sample_ID, Abs_562nm_raw, Abs_562nm_corrected, Dilution_factor, Volume_ul, TotalProtein_ug, TotalProtein_ug_mL)

# View(TotalProtein_final)
nrow(TotalProtein_final) # 118

# NOTE! I found that some of my discriminants are negative (b^2 - 4ac) IN PLATE 2!! (as 'NaN)
# so Im going to extrapolate from a linear regression curve based on the final samples 

# first lets plot absorbance and actual concentration calacuated 
calc_BCA_plot <- TotalProtein_final %>% 
                    ggplot(aes(y = Abs_562nm_corrected, 
                               x  = TotalProtein_ug_mL)) +
                          geom_point() + 
                    theme_bw() + 
                    ggtitle('Total protein: Calculated BCA by Net Absorbance') +
                    facet_wrap(~Plate)
calc_BCA_plot 

#print
pdf(paste("Output/catalase/plots/Total_Protein_standard_curves_with_data.pdf", sep =''), 
    width=10, 
    height=8)
ggpubr::ggarrange(BCA_stand_plots_quadratic,
          calc_BCA_plot,
          nrow = 2)
dev.off()
# write csv
write.csv(TotalProtein_final, file = "Output/catalase/Total_protein_master.csv")


```


## CATALASE 

* Generated curve with standards
  - calculate the mean blank standard as 0 catalase_UmL or Standards 'S7'
  - correct standards for the blank absorbance 
  
  
* Calcaulate the sample U/mL 
  - correct absorbance for blank 
  - use quadratic equation from standard curve to calculate U/mL catalase


* STEP 1: Calculate blank absorbance

```{r catalase blank absorbnace}

# Plate 1 means - then calc DA412 = (A412)final – (A412)initial
Plate1.blanks        <- catalase_plate1.standards %>% dplyr::filter(catalase_UmL  == 0)
Plate1.blank.mean    <- mean(Plate1.blanks$Cat_Absorbance_560nm) # 1.18315


# Plate 2 means - then calc DA412 = (A412)final – (A412)initial
Plate2.blanks        <- catalase_plate2.standards %>% dplyr::filter(catalase_UmL  == 0)
Plate2.blank.mean    <- mean(Plate2.blanks$Cat_Absorbance_560nm) # 1.0943

```

* STEP 2: Generate standard curve (correct for blank background abs first!)

  - plate1: y = SSasymp(x, 0.34, 7.30, 2.14); x = Abs at 560 um, y = catalase U mL
  - plate2: y = SSasymp(x, 0.10, 6.35, 1.49); x = Abs at 560 um, y = catalase U mL
  
  * **Important** - 25 ul of sample was measured without dilution for catalase. This is critical to correct to total protein
  as our protocol for total protein also used 25 ul per sample - so it will be 1:1 to correct for catalase per total protein
  
```{r generate standard curve}
library(drc) # install.packages('drc')
## Objectve to subtract the blank from all standards and samples. and build standard curve with equation

# Plate 1
Plate1_standards_meanSE <- catalase_plate1.standards %>% 
                                    dplyr::rename(Absorbance_560nm_raw = Cat_Absorbance_560nm) %>% 
                                    dplyr::select(Plate, Absorbance_560nm_raw, catalase_UmL) %>% 
                                    dplyr::group_by(Plate, catalase_UmL) %>% 
                                    dplyr::summarise(meanAbs = mean(Absorbance_560nm_raw),
                                                     sdAbs   = sd(Absorbance_560nm_raw),
                                                     n = n())
Plate1_standards_model <- drm(catalase_UmL ~ meanAbs, data = Plate1_standards_meanSE, fct = LL.4()) # generate four parametr logistic model
newdata_plate1         <- data.frame(meanAbs = seq(min(Plate1_standards_meanSE$meanAbs),
                                                   max(Plate1_standards_meanSE$meanAbs), length.out = 100))
predictions_plate1     <- predict(Plate1_standards_model, newdata = newdata_plate1)
params_plate1          <- coef(Plate1_standards_model) # # Extract parameters

# equation_plate1        <- sprintf("y = %.2f + (%.2f - %.2f) / (1 + (x / %.2f)^%.2f)",
#                                   coef(Plate1_standards_model)[1],
#                                   coef(Plate1_standards_model)[2],
#                                   coef(Plate1_standards_model)[1],
#                                   coef(Plate1_standards_model)[3],
#                                   coef(Plate1_standards_model)[4])


Plate1_fit <- nls(catalase_UmL ~ SSasymp(meanAbs, yf, y0, log_alpha), data = Plate1_standards_meanSE)
summary(Plate1_fit)
# Formula: y ~ SSasymp(x, yf, y0, log_alpha)
# Parameters:
#           Estimate Std. Error t value Pr(>|t|)   
# yf          0.3360     0.1966   1.709  0.16266   
# y0          7.3016     1.1811   6.182  0.00348 **
# log_alpha   2.1397     0.2744   7.798  0.00146 **
equation_plate1        <- sprintf("y = SSasymp(x, %.2f, %.2f, %.2f)",
                                  summary(Plate1_fit)$coefficients[1], # yf
                                  summary(Plate1_fit)$coefficients[2], # y0
                                  summary(Plate1_fit)$coefficients[3] # log_alpha
                                  )




# Plate1_model <- lm(Plate1_standards_meanSE$catalase_UmL ~ poly( Plate1_standards_meanSE$meanAbs, 4))
# summary(Plate1_model)$coefficients
#                                            Estimate Std. Error   t value   Pr(>|t|)
# (Intercept)                                1.406286  0.1920777  7.321443 0.01814917
# poly(Plate1_standards_meanSE$meanAbs, 4)1 -3.773888  0.5081897 -7.426139 0.01765440
# poly(Plate1_standards_meanSE$meanAbs, 4)2  1.681861  0.5081897  3.309513 0.08043806
# poly(Plate1_standards_meanSE$meanAbs, 4)3 -1.098399  0.5081897 -2.161395 0.16320663
# poly(Plate1_standards_meanSE$meanAbs, 4)4  0.829503  0.5081897  1.632270 0.24421454

# y = e + dx + cx^2 + bx^3 + ax^4

# equation_plate1        <- sprintf("y = %.2f + %.2f*x + %.2f*x^2 + %.2f*x^3 + %.2f*x^4",
#                                   summary(Plate1_model)$coefficients[5], # e
#                                   summary(Plate1_model)$coefficients[4], # d
#                                   summary(Plate1_model)$coefficients[3], # c
#                                   summary(Plate1_model)$coefficients[2], # b
#                                   summary(Plate1_model)$coefficients[1]  # a
                                  # )



Plate1_Catalase_curve <- ggplot(Plate1_standards_meanSE, aes(x = meanAbs, y = catalase_UmL)) + # # Plot the data and fitted curve
                                  geom_point() +
                                  geom_line(data = data.frame(meanAbs = 
                                                                newdata_plate1$meanAbs, response = predictions_plate1), 
                                            aes(x = meanAbs, y = response), color = "blue", size = 1) +
                                  labs(x = "meanAbs", y = "catalase_UmL", title = "Catalase curve Plate 1") +
                                  annotate("text", x = Inf, y = Inf, label = equation_plate1, hjust = 1, vjust = 1) +
                                  theme_minimal()

# y = SSasymp(x, 0.34, 7.30, 2.14)



# plate 2


Plate2_standards_meanSE <- catalase_plate2.standards %>% 
                                    dplyr::rename(Absorbance_560nm_raw = Cat_Absorbance_560nm) %>% 
                                    dplyr::select(Plate, Absorbance_560nm_raw, catalase_UmL) %>% 
                                    dplyr::group_by(Plate, catalase_UmL) %>% 
                                    dplyr::summarise(meanAbs = mean(Absorbance_560nm_raw),
                                                     sdAbs   = sd(Absorbance_560nm_raw),
                                                     n = n())
Plate2_standards_model <- drm(catalase_UmL ~ meanAbs, data = Plate2_standards_meanSE, fct = LL.4()) # generate four parametr logistic model
newdata_plate2         <- data.frame(meanAbs = seq(min(Plate2_standards_meanSE$meanAbs), max(Plate2_standards_meanSE$meanAbs), length.out = 100))
predictions_plate2     <- predict(Plate2_standards_model, newdata = newdata_plate2)
params_plate2          <- coef(Plate2_standards_model) # # Extract parameters
equation_plate2        <- sprintf("y = %.2f + (%.2f - %.2f) / (1 + (x / %.2f)^%.2f)",
                                  params_plate2[1], params_plate2[2], params_plate2[1], params_plate2[3], params_plate2[4])




Plate2_fit <- nls(catalase_UmL ~ SSasymp(meanAbs, yf, y0, log_alpha), data = Plate2_standards_meanSE)
summary(Plate2_fit)
# Formula: catalase_UmL ~ SSasymp(meanAbs, yf, y0, log_alpha)
# Parameters:
#           Estimate Std. Error t value Pr(>|t|)    
# yf         0.09717    0.09858   0.986     0.38    
# y0         6.34540    0.25645  24.743 1.58e-05 ***
# log_alpha  1.48719    0.08554  17.385 6.43e-05 ***
equation_plate2        <- sprintf("y = SSasymp(x, %.2f, %.2f, %.2f)",
                                  summary(Plate2_fit)$coefficients[1], # yf
                                  summary(Plate2_fit)$coefficients[2], # y0
                                  summary(Plate2_fit)$coefficients[3] # log_alpha
                                  )
# y = SSasymp(x, 0.10, 6.35, 1.49)

# Plate2_model <- lm(Plate2_standards_meanSE$catalase_UmL ~ poly( Plate2_standards_meanSE$meanAbs, 4))
# summary(Plate2_model)$coefficients
#                                             Estimate Std. Error    t value     Pr(>|t|)
# (Intercept)                                1.4062857 0.04003263  35.128491 0.0008093819
# poly(Plate2_standards_meanSE$meanAbs, 4)1 -3.9189487 0.10591637 -37.000405 0.0007296448
# poly(Plate2_standards_meanSE$meanAbs, 4)2  1.7697783 0.10591637  16.709205 0.0035625650
# poly(Plate2_standards_meanSE$meanAbs, 4)3 -0.9601472 0.10591637  -9.065144 0.0119511667
# poly(Plate2_standards_meanSE$meanAbs, 4)4  0.2176236 0.10591637   2.054674 0.1762623456

# y = e + dx + cx^2 + bx^3 + ax^4

# equation_plate2        <- sprintf("y = %.2f + %.2f*x + %.2f*x^2 + %.2f*x^3 + %.2f*x^4",
#                                   summary(Plate2_model)$coefficients[5], # e
#                                   summary(Plate2_model)$coefficients[4], # d
#                                   summary(Plate2_model)$coefficients[3], # c
#                                   summary(Plate2_model)$coefficients[2], # b
#                                   summary(Plate2_model)$coefficients[1]  # a
#                                   )

# equation_plate1        <- sprintf("y = %.2f + %.2f*x + %.2f*x^2 + %.2f*x^3 + %.2f*x^4",
#                                   summary(Plate1_model)$coefficients[5], # e
#                                   summary(Plate1_model)$coefficients[4], # d
#                                   summary(Plate1_model)$coefficients[3], # c
#                                   summary(Plate1_model)$coefficients[2], # b
#                                   summary(Plate1_model)$coefficients[1]  # a
                                  
Plate2_Catalase_curve <- ggplot(Plate2_standards_meanSE, aes(x = meanAbs, y = catalase_UmL)) + # # Plot the data and fitted curve
                                  geom_point() +
                                  geom_line(data = data.frame(meanAbs = 
                                                                newdata_plate1$meanAbs, response = predictions_plate1), 
                                            aes(x = meanAbs, y = response), color = "blue", size = 1) +
                                  labs(x = "meanAbs", y = "catalase_UmL", title = "Catalase curve Plate 2") +
                                  annotate("text", x = Inf, y = Inf, label = equation_plate2, hjust = 1, vjust = 1) +
                                  theme_minimal()

# y = 0.22 + -0.96*x + 1.77*x^2 + -3.92*x^3 + 1.41*x^4

# output 
pdf(paste("Output/catalase/plots/CatalaseUmL_standard_curves.pdf", sep =''), 
    width=10, 
    height=6)
print(ggarrange(Plate1_Catalase_curve,Plate2_Catalase_curve, nrow = 1))
dev.off()





```


## Calculate catalase


* Steps: 

  - assemble the averages for total protein (duplicates) as 'TotalProtein_final_MeanSE'
  
  - merge means total protein with the catalase absorbance data
  
    - quick sanity check to ensure the merge was successful
    
  - calculate catalase using the standard equations from the chunk above

```{r average total protein by sample id}
library(tidyr)

# first prepare total protein to merge with the catalase samples
# we want ot take a mean for the total protein value since we measured in duplicate
# so this below, has 59 unique Ids, this is correct
TotalProtein_final_MeanSE <- TotalProtein_final %>% 
                                 dplyr::select(Plate, Sample_ID, TotalProtein_ug) %>% 
                                 Rmisc::summarySE(measurevar = 'TotalProtein_ug', groupvars = 'Sample_ID') %>% 
                                 dplyr::select(Sample_ID, TotalProtein_ug)
nrow(TotalProtein_final_MeanSE) # 58 unique IDs , took the mean for total protein between replicates!

```

```{r merge total protein wit caltalase samples}
# now bind together all the catalase samples
catalase_all <- rbind(catalase_plate1.samples, catalase_plate2.samples)
nrow(catalase_all) # 120
length(unique(catalase_all$Sample_ID)) # 60 unique sample ids, correct good to go

# merge with total protein
catalase_all_protein <- merge(catalase_all, # merge calatase all
                              TotalProtein_final_MeanSE, # with three columns of total protein final
                              by = "Sample_ID") # by these ccommon columns
nrow(catalase_all_protein) # 116
length(unique(catalase_all_protein$Sample_ID)) # 58, which samples did we lose?

```


# Calculate catalase 

  * calculation uses the equations shown for the standards in chunks above
  
    - before proceeding (to final data and plots) plot a snaity check of standard curves to calculated values by absorbance 
    
  * final master dataset of means by sample ID
  
  * graph the data as mean SE
  
```{r calculate catalase}


Catalase_Uml_ngprotein_final <- catalase_all_protein %>% # row bind, merge both together
                                          dplyr::mutate(
                                            Catalase_UmL = # new column
                                              case_when( # dependencies of the new column  when...
                                                Plate %in% 1 ~ SSasymp(Cat_Absorbance_560nm, 0.34, 7.30, 2.14),
                                                Plate %in% 2 ~ SSasymp(Cat_Absorbance_560nm, 0.10, 6.35, 1.49)
                                                #Plate %in% 1 ~ 0.82 + (-0.34 - 0.82) / (1 + (Cat_Absorbance_560nm / 24.77)^0.01), # column plate is 1, EQ,
                                                #Plate %in% 2 ~ 1.66 + (-0.22 - 1.66) / (1 + (Cat_Absorbance_560nm / 5.63)^0.20) # column plate is 2, EQ,
                                              ),
                                            Catalase_UmL_ng_protein = 
                                              Catalase_UmL / (TotalProtein_ug /1000),
                                            Treatment = 
                                              case_when(
                                                grepl("BL", Sample_ID) ~ "Baseline", 
                                                grepl("A1|A2|A3|B1|B2|B3", Sample_ID) ~ "Control",
                                                grepl("A4|A5|A6|B4|B5|B6", Sample_ID) ~ "OA",
                                                grepl("C1|C2|C3|D1|D2|D3", Sample_ID) ~ "HW"),
                                            Time = 
                                              case_when(
                                                grepl("BL", Sample_ID) ~ "Baseline", 
                                                grepl("T1", Sample_ID) ~ "1",
                                                grepl("T2", Sample_ID) ~ "2",
                                                grepl("T3", Sample_ID) ~ "3")
                                              ) %>% 
                                           dplyr::select(Plate, Sample_ID, Treatment, Time, 
                                                         Cat_Absorbance_560nm, TotalProtein_ug, 
                                                         Catalase_UmL, Catalase_UmL_ng_protein)
                            


# Sanity check figures of standards and calculated values (not corrected for total protein)
calc_Catalase_plot <- Catalase_Uml_ngprotein_final %>% 
                    ggplot(aes(y = Catalase_UmL, 
                               x  = Cat_Absorbance_560nm)) +
                          geom_point() + 
                    theme_bw() + 
                    ggtitle('Catalase U mL: Calculated by Absorbance') +
                    facet_wrap(~Plate, scales  = 'free_y')

Standard_curves_with_Catalasedata <-  ggpubr::ggarrange(
                                                  ggpubr::ggarrange(Plate1_Catalase_curve, Plate2_Catalase_curve, nrow =1),
                                                  calc_Catalase_plot, nrow = 2)

pdf(paste("Output/catalase/plots/CatalaseUmL_standard_curves_with_data.pdf", sep =''), 
    width=10, 
    height=6)
print(Standard_curves_with_Catalasedata)
dev.off()

# final dataset of means by sample ID 


Catalase_Uml_ngprotein_final_MeanSE <- Catalase_Uml_ngprotein_final %>% Rmisc::summarySE(groupvars = c('Plate','Treatment','Time', 'Sample_ID'), measurevar = 'Catalase_UmL_ng_protein')


# plot Catalase mean SE


pd <- position_dodge2(width = 0.2)

Catalase_MeanSE <- Catalase_Uml_ngprotein_final_MeanSE %>% 
  
                      group_by(Treatment, Time) %>% 
                      
                       dplyr::summarise( # summarise to aquire the mean and SE for plotting
                        Catalase_mean = mean(Catalase_UmL_ng_protein), # mean
                        Catalase_sd = sd(Catalase_UmL_ng_protein), # sd
                        n = n(), # count
                        Catalase_se = Catalase_sd / sqrt(n)) %>% # SE
                      
                      # plot it
                      ggplot(aes(x=Time, y=Catalase_mean, group=Treatment)) + 
                        geom_line(aes(group = factor(Treatment)), size = 0.5, position=position_dodge(.4)) +  # connect a line between variables
                        # scale_linetype_manual(values=c("solid", "dashed", "dotted")) +
                        geom_point(aes(shape=Treatment, fill=Treatment), size = 4.5,position=position_dodge(.4)) + 
                        # scale_shape_manual(values=c(21, 22, 24)) + # filled circle, filled triangle, and X 
                        # scale_fill_manual(values=c("#009E73","#E69F00", "#CC79A7")) + # fill cicle white, triangle white, and jsut black for the X
                        geom_errorbar(aes(ymin=(Catalase_mean)-(Catalase_se), 
                                          ymax=(Catalase_mean)+(Catalase_se)), 
                                      width=0,position=position_dodge(.4)) + # width dtermines the length of the end ticks
                        # geom_jitter() +
                        theme_classic() +
                        theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank()) +#,
                              # legend.position = "none") + 
                        ggtitle("Catalase (UmL ng protein-1)") +
                        labs(y= "Catalase_UmL_ng_protein", x = "Time")


 # output files 
pdf(paste("Output/catalase/plots/CatalaseUmL_ng_protein.pdf", sep =''), 
    width=8, 
    height=6)
print(Catalase_MeanSE)
dev.off()

# write csv
write.csv(Catalase_Uml_ngprotein_final, file = "Output/catalase/Catalase_calculated_master.csv")
write.csv(Catalase_Uml_ngprotein_final_MeanSE, file = "Output/catalase/Catalase_calculated_masterMeanSE.csv")
 
```
